---
title: "Put a real title here"
subtitle: "SDS 192: MP4"
author: 
- name: Emily Rhyu
  affiliation: Smith College
- name: Alicia Chen
  affiliation: Smith College
- name: Elda Cervantes
  affiliation: Smith College

date: "`r format(Sys.Date(), '%B %e, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 8
    fig_height: 8
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE,
  warning = FALSE
)
library(tidyverse)
library(sds192)
library(sf)
library(tidygeocoder)
```

## Introduction

> READ THE INSTRUCTIONS: https://beanumber.github.io/sds192/mod_spatial.html

Biking from Smith College to MacLeish is of interest to a subset of field station visitors. Considering that some bikers appreciate climbing hills and others wish to avoid hills, propose two alternative bike routes to the field station using the available data from MassGIS. Detail the lengths of the routes and the elevation change along the routes, and support your choices with data. It would be cool if you could create elevation profiles (e.g., this one) for each bike route.
Supporting data includes:
- Road shapefiles (retrieve from MassGIS)
- 30’ contour elevation data (elevation)
- Property boundary layer (boundary)


## Loading the data

```{r, message=FALSE}
library(macleish)
names(macleish_layers)
```

```{r}
points_of_interest <- tribble(
  ~name, ~lat, ~long,
  "Group Campsite", 42.450976, -72.678154, 
  "Remote campsite", 42.458549, -72.679581,
  "Parking", 42.448013, -72.680663,
  "Weather Station OrchardMet", 42.449653, -72.680315,
  "Weather Station WhatelyMet", 42.448470, -72.680553
) %>%
  st_as_sf(coords = c("long", "lat")) %>%
  st_set_crs(4326)
```

```{r}
trails <- mass_gis("dcrtrails")

boundary <- macleish_layers %>%
  pluck("boundary")

elevations <- macleish_layers %>%
  pluck("elevation")
```

```{r}
locations <- tribble(
  ~place, ~address, 
  "Smith", "44 College Lane, Northampton, MA 01063",
  "Macleish", "30, 78 Poplar Hill Rd, South Deerfield, MA 01373"
  
) %>%
  #geocoding
  geocode(address, method = "osm") %>%
  st_as_sf(coords = c("long", "lat")) %>%
  st_set_crs(4326)

#maclongitude="-72.68171351473106"
#maclatitude="42.45467547936159"
#macleish <- tibble(place = "Macleish", #geometry="-72.68171351473106,42.45467547936159"
#geometry=st_sfc(st_point(c(42.45467547936159,-72.68171351473106)))
#)

#locations <- locations %>% 
#  inner_join(macleish, by = c("geometry"="geometry"))
```

```{r, eval=FALSE}
#location of macleish
mac <- leaflet() %>%
  addTiles() %>%
  addMarkers(
    lng =  -72.68171351473106,
    lat = 42.45467547936159,
    popup = "Macleish"
  )
mac
```


```{r, eval=FALSE}
#might use to find flat/hilly path (50 is flattest/easiest)
#use summarize?
rating = sqrt(2 x gain x distance)

#finds elevation for trail in question. probably use to apply to flat/hilly trails found from equation above and make elevation profile
trail_elevations <- longest_trail %>%
  st_intersection(elevations) %>%
  st_cast("MULTIPOINT") %>%
  st_cast("POINT")
```



## Static maps

You can make a basic static map with **ggplot2** using `geom_sf()`, as shown in the figure below.

```{r static, fig.cap="The boundary of the MacLeish property.",eval=FALSE}

ggplot() +
  geom_sf(data = boundary) +
  geom_sf(data = elevations)
```

```{r, eval=FALSE}
#example we can reference
boundary_plot +
  geom_sf(data = elevations, color = "dark gray") + 
  geom_sf(data = longest_trail, color = "brown", size = 1.5) + 
  geom_sf(data = trail_elevations, fill = "yellow", pch = 21, size = 3) + 
  geom_sf_label(
    data = trail_elevations, 
    aes(label = CONTOUR_FT), 
    hjust = "right", 
    size = 2.5,
    nudge_x = -0.0005
  )
```


## Dynamic maps

You can make a basic dynamic map with **leaflet**, as shown in the figure below.

```{r dynamic, fig.cap="The boundary of the MacLeish property.",eval=FALSE}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addPolygons(data = boundary)
```



## Word count

```{r word_count, message=FALSE, echo=FALSE}
text_stats()
```

## Standards

In this assignment, we attempted the following [standards](https://beanumber.github.io/sds192/standards.html):

- `r standard_button("wrangling")`: We met the Wrangling standard because we...
- `r standard_button("spatial")`: We met the Spatial standard because we...
- `r standard_button("leaflet")`: We met the Leaflet standard because we...
- `r standard_button("aesthetics")`: We mastered the Aesthetics standard because we...
- `r standard_button("markdown")`: We mastered the R Markdown standard because we...

## References
